---
title: 'SMARTDADA2: Interactive Output'
output: html_document
params:
  trim_file: "file_path"
  sumEE_file: "file_path"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(include = FALSE)
knitr::opts_chunk$set(warning = FALSE)

# Check dependecies
packages <- c("readr", "dplyr", "reshape2", "ggplot2", "argparse", "plotly", "cowplot", "rmarkdown")

for (package in packages) {
  if(!require(package, character.only = T)){
    install.packages(package)
    library(package)
  }
}
```

```{r}
sumEE_info <- read_delim(paste0("../../" ,sumEE_file),
                         delim = "\t", escape_double = FALSE,
                         trim_ws = TRUE)
trim_info <- read_delim(paste0("../../" ,trim_file),
                        delim = "\t", escape_double = FALSE,
                        col_types = cols(LeftIndex = col_integer(),
                                         RightIndex = col_integer(),
                                         ReadLength = col_integer()),
                        trim_ws = TRUE)
```

```{r}
lefttrim <- unique(trim_info$LeftTrim)
righttrim <- unique(trim_info$RightTrim)

HeatMap <- ggplot(trim_info, aes(LeftIndex, RightIndex, fill= AvgEEPerPosition)) +
            geom_tile() +
            geom_vline(xintercept = lefttrim) +
            geom_hline(yintercept = righttrim)+
            scale_y_continuous(trans = "reverse", breaks = unique(trim_info$RightIndex)) +
            scale_fill_gradient(low="white", high="blue") +
            theme_bw() +
            theme(plot.title = element_text(hjust = 0.5),
                  text = element_text(size = 14),
                  axis.text.x = element_text(size = 10),
                  axis.text.y = element_text(size = 10))

ScatterPlot <- ggplot(trim_info, aes(ReadLength, AvgEEPerPosition)) +
                geom_point() +
                theme_bw() +
                theme(text = element_text(size = 14)) +
                ylab('Average Expected Error Per Position') +
                xlab('Read Length (bp)')

melt_sumEEInfo <- melt(sumEE_info)
Histogram <- ggplot(melt_sumEEInfo, aes(x = value, fill = variable)) +
              geom_histogram(position = "dodge") +
              geom_vline(xintercept = 2) +
              theme_bw() +
              theme(text = element_text(size = 14)) +
              scale_x_continuous(limits = c(0,4)) +
              guides(fill=guide_legend(title=" "))+
              xlab('Sum of Expected Error')

h_hm <- length(unique(trim_info$RightIndex))*15

pHeatMap <- ggplotly(HeatMap, height = h_hm)
pScatterPlot <- ggplotly(ScatterPlot)
pHistogram <- ggplotly(Histogram)
```

### Understanding the effect of trimming and truncating on quality scores
- vertical line represents the detected obvious trimming parameter **(e.g, trim-left)**
- horizontal line represents the detected obvious truncation parameter **(e.g., trunc-len)**

Note: If lines are at zero no obvious trimming or truncating position was found

```{r, include=TRUE}
pHeatMap
```

### Understanding the effect of read length on quality scores

```{r, include=TRUE}
pScatterPlot
```

### Understanding the number of reads retained with obvious trimming

- $\color{blue}{\text{Blue}}$ lines are counts of reads with a certain quality score when obvious trimming has been performed
- $\color{red}{\text{Red}}$ lines are from reads where no trimming has been performed
- Vertical line: DADA2 defaults **max-ee** parameter to 2

```{r, include=TRUE}
pHistogram
```

